---
title: "Getting started"
output:
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
your_local_path <- '/Users/dennisbrookner/github'
```
## Installation
Install via:
```{r, echo=FALSE, message=FALSE, results='hide'}
install.packages(c("devtools", "tidyverse"), repos = "http://cran.us.r-project.org")
```
```{r, message=FALSE, results='hide'}
devtools::install_github("Puntalytics/puntr")
library(puntr)
library(tidyverse) # always a good idea to do this too
```
## Using `puntr` with NFL data
### Completed seasons
For speed, we've already scraped (using [`nflfastR`](https://www.nflfastr.com/)) and saved punting data for the 1999-2020 seasons. The easiest thing to do is download the `puntr-data` repo [here](https://github.com/Puntalytics/puntr-data/tree/master/data), and then point `puntr::import_punts()` to your local copy of the data. You can also download the data directly each time; this takes around 15 minutes.  
Import, clean, and calculate as follows:
```{r imports, message=FALSE}
#punts_raw <- import_punts(1999:2020, local=TRUE, path=your_local_path) # recommended
punts_raw <- import_punts(2018:2020) # This takes ~15 minutes
punts_cleaned <- trust_the_process(punts_raw) # clean
punts <- calculate_all(punts_cleaned) # calculate custom Puntalytics metrics 
```
You now have a dataframe `punts` where each row is a punt, and each column is a stat relevant to punting (including our custom metrics).

### In-progress seasons
The kind folks at [`nflfastR`](https://www.nflfastr.com/) have set up a convenient `SQL`-y way to scrape data for in-progress seasons. Rather than redo all of that work, we'll just share here the code we use for this purpose:
```{r eval=FALSE}
#install.packages("DBI")
#install.packages("RSQLite")
library(DBI)
library(RSQLite)
library(nflfastR)
library(puntr)

update_db()
connection <- dbConnect(SQLite(), "./pbp_db")
pbp <- tbl(connection, "nflfastR_pbp")

punts <- pbp %>% filter(punt_attempt==1) %>%
  collect() %>%
  trust_the_process() %>%
  calculate_all()

dbDisconnect(connection)
```

### Comparing punters
To compare punters, continue with
```{r}
mini <- create_mini(punts)
```
to get a dataframe where each row is a punter, and each column is an average stat for that punter. There are many columns in this dataframe, but here's a few of the most interesting:  
```{r}
mini <- mini %>% arrange(desc(Punt_eaepaae_avg)) %>% select(punter_player_name, Gross, Net, SHARP_RERUN, Punt_eaepaae_avg)
rmarkdown::paged_table(mini)
```
To compare punter **seasons**, instead use
```{r}
miniY <- create_miniY(punts)
```
which gives every unique punter season a row.  
```{r, echo=FALSE}
miniY <- miniY %>% arrange(desc(Punt_eaepaae_avg))
rmarkdown::paged_table(miniY)
```
To compare punter **games**, instead use
```{r}
miniG <- create_miniG(punts)
```
which gives every unique punter game a row.  
```{r, echo=FALSE}
miniG <- miniG %>% arrange(desc(Punt_eaepaae_avg))
rmarkdown::paged_table(miniG)
```
  
These dataframes - `punts`, `mini`, `miniY` and `miniG` - should serve as a good starting point for any custom analysis you'd like to do, be that using built-in `puntr` metrics, or your own.

## Using `puntr` with college data
***NOTE: `puntr` was successfully migrated from `cfbscrapR` to `cfbfastR` in version 1.2.2***
  
`puntr` can also handle punting data for college football, piggybacking off of the scraping abilities of the [`cfbfastR`](https://saiemgilani.github.io/cfbfastR/) package. Import and clean as follows:
```{r message=FALSE, warning=FALSE, eval=FALSE}
college_punts <- import_college_punts(2019:2020) %>% # import (calls cfbfastR behind the scenes)
 college_to_pro() %>% # rename columns to those used by nflfastR
 calculate_all() # calculate as with NFL data
```
Now we can use the same `create_miniY` as above to compare college punter seasons (`create_mini` and `create_miniG` would also work here, of course.)
```{r, eval=FALSE}
miniY_college <- create_miniY(college_punts)
```
```{r, echo=FALSE, eval=FALSE}
miniY_college <- miniY_college %>% arrange(desc(SHARP_RERUN))
rmarkdown::paged_table(miniY_college)
```
