---
title: "Getting started"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## puntr
Install via:
```{r, message=FALSE}
devtools::install_github("Puntalytics/puntr")
library(puntr)
library(tidyverse)
```
Once that's installed, you can import, clean, and calculate some punting data:
```{r}
# punts_raw <- import_punts(1999:2019) # this takes a minute or two - there have been a lot of punts since 1999!
punts_raw <- import_punts(2018:2019) # fewer seasons is faster
punts_cleaned <- trust_the_process(punts_raw) # process
punts <- calculate_all(punts_cleaned) # calculate custom Puntalytics metrics 
```
You now have a dataframe `punts` where each row is a punt, and each column is a stat relevant to punting (including our custom metrics).  
  
To compare punters, continue with
```{r}
mini <- create_mini(punts)
```
to get a dataframe where each row is a punter, and each column is an average stat for that punter.  
```{r, echo=FALSE}
mini <- mini %>% arrange(desc(SHARP_RERUN))
rmarkdown::paged_table(mini)
```

To compare punter **seasons**, instead use
```{r}
miniY <- create_miniY(punts)
```
which gives every unique punter season a row.  
```{r, echo=FALSE}
miniY <- miniY %>% arrange(desc(SHARP_RERUN))
rmarkdown::paged_table(miniY)
```
  
These three dataframes - `punts`, `mini`, and `miniY` - should serve as a good starting point for any custom analysis you'd like to do, be that using built-in `puntr` metrics, or your own.
